---
title: "Assessment1"
author: "Michael Yefremov"
date: "Sunday, April 19, 2015"
output: html_document
---
###Loading and preprocessing data
The first step of this assigment is to load data. So be sure that there are unziped file with data ("activity.csv") in your working directory.
```{r load data}
setwd("e:/coursera/Reproduc research") #setting work directory
fl <- read.csv ("./activity.csv", na.string = "NA")
str(fl) #Look briefly on data
```
So there are 17 568 observations and 3 variable. **steps** and **interval** has an integer type and **date** is as factor.  
Let converts variable **date** into Date-format for convinience
```{r convert type}
fl$date <- as.Date (fl$date)
```

###What is mean total number of steps taken per day?

Now it should be obtain the total number of steps per day. For this I will use *plyr* package. But at first exclude NA values.

```{r obtaining total steps per day}
good <- complete.cases (fl)
fl_g <- fl[good,] #variable fl_g is without NA values

library(plyr)
StepsPerDay <- ddply (fl_g, .(date), summarise, steps_pd = sum(steps))
```

So we can make a histogram of the total number of steps taken each day. I will use **ggplot2** graphic system

```{r make hist}
library (ggplot2)
ggplot(StepsPerDay, aes(date, steps_pd)) + geom_bar (stat = "identity") + labs(title = "Total number of steps each day", x = "Date", y = "Steps")
```

Empty spaces on the plot mean that in this day there were no steps or we have no data.  
Differences between bar charts and histogram: With bar charts, each column represents a group defined by a categorical variable; and with histograms, each column represents a group defined by a quantitative variable. In our case there is histogram because each column represents number of steps (quantitaive variable).  
Now I'm going to calculate **mean** and **median** of the total number of the steps per day. 
```{r calculating mean and meadian}
mean_withoutNA <- ceiling(mean (StepsPerDay$steps_pd))
median_withoutNA <- median (StepsPerDay$steps_pd)
```
So **mean** is `r as.integer(mean_withoutNA)` and **median** is `r median_withoutNA`.

###What is the average daily activity pattern?
Let it obtains average daily activity and make a time series plot.

```{r daily }
DailyActivity <- ddply (fl_g, .(interval), summarise, steps_pi = mean(steps))

ggplot(DailyActivity, aes(interval, steps_pi)) + geom_line (size = 1) + labs(title = "Average daily activity", x = "Interval", y = "Steps")
```

Now I am going to obtain 5-minute interval in which average number of steps is maximum.

```{r obtaining interval with maximum steps}
max_i <- DailyActivity [DailyActivity$steps_pi == max(DailyActivity$steps_pi), "interval"]
```

So the 5-minute interval which contains the maximum number of steps is **`r max_i`**

###Inputing missing values

Now I'm going to obtain number of missing values:
```{r number of missing values}
sum (!good) #variable good contain number of complete cases in variable with data
```

I've decided use mean for 5-minute interval with NA values. Now I'm going to create a new data set with filled missing NA

```{r filling msiing NA}
fl_n <- fl #creating new data frame
for (i in 1:length (fl_n$steps)){
        if (is.na(fl_n[i, "steps"])) {
                gd <- fl_n[i,"interval"] == DailyActivity$interval
                fl_n[i,"steps"] <- ceiling(DailyActivity[gd, "steps_pi"])
                }
}
```

Now I'm going to make a histogram of a new variable without NA
```{r making hist of new data set total number of steps taken each day}
StepsPerDay_n <- ddply (fl_n, .(date), summarise, steps_pd = sum(steps))

ggplot(StepsPerDay_n, aes(date, steps_pd)) + geom_bar (stat = "identity") + labs(title = "Total number of steps each day without NA", x = "Date", y = "Steps")
```

Now I'm going to calculate **NEW** **mean** and **median** of the total number of the steps per day and compare they with previous. 
```{r calculating NEW mean and meadian}
mean_filledNA <- ceiling(mean (StepsPerDay_n$steps_pd))
median_filledNA <- median (StepsPerDay_n$steps_pd)
```

So NEW **mean** is `r as.integer(mean_filledNA)` and is PREVIOUS (where NA omit) `r as.integer(mean_withoutNA)`. So adding filled NA values slightly increases **mean**.  
NEW **median** is `r as.integer(median_filledNA)` and is PREVIOUS (where NA omit) `r as.integer(median_withoutNA)`. So adding filled NA values increases **median**.  

###Are there differences in activity patterns between weekdays and weekends?
Now I'm going to create a new factor variable with 2 levels Weekend & weekday

```{r making factor variable}
fl_n$weekdays <- weekdays(fl_n$date)
we <- fl_n$weekdays %in% c("Sunday", "Saturday")
fl_n$weekdays[we] <- "Weekend"
fl_n$weekdays[!we] <- "Weekday"
fl_n$weekdays <- factor(fl_n$weekdays)
```

Now I'm going to calculate 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend. And then plot it

```{r making 5 min intervals}
DailyActivity_w <- ddply (fl_n, .(interval, weekdays), summarise, steps_pi = mean(steps))

ggplot(DailyActivity_w, aes(interval, steps_pi)) + geom_line (size = 1) + labs(title = "Average daily activity", x = "Interval", y = "Steps") + facet_grid(weekdays~.)
```

As we can see there are differences in activity patterns between weekdays and weekends